Near and Mid-infrared properties of known $z\geq5$ Quasars

### 2019-06-05 tex/VHzQ.tex
Corrections to English in Section 3.1

We use data from the the beginning of the WISE mission \citep[2010 January; ][]{Wright2010} through the fifth-year of NEOWISE-R operations \citep[2018 December;]{Mainzer2011}.  There are several  major \href{https://irsa.ipac.caltech.edu/Missions/wise.html}{data releases} and catalogues based on the WISE mission. Here we use two: the WISE \href{http://wise2.ipac.caltech.edu/docs/release/allwise/}{AllWISE Data release} and the recently released ``unWISE Catalog'' \citet{Schlafly2019}. The AllWISE program combines the W1 and W2 Single-exposure data from all WISE survey phases (4-Band Cryo, 3-Band Cryo and Post-Cryo; 2010-01-07 thru 2011-02-01) survey phases, and the W3 and W4 from the 4-Band Cryo phase. The unWISE effort\footnote{\href{http://unwise.me}{http://unwise.me}} is the unblurred coadds of the WISE imaging using the AllWISE and NEOWISE-R stacked data \citep{Lang2014, Meisner2018a, Meisner2018b}.

For the two shorter WISE bands, $\lambda_{\rm eff}$3.37$\mu$m W1 and $\lambda_{\rm eff}$4.62$\mu$m W2 we generally report the deeper, unblurred unWISE coadd data.  For the two longer WISE bands, $\lambda_{\rm eff}$12.1$\mu$m and $\lambda_{\rm eff}$22.8$\mu$m we use the ALLWISE Data Release.

Previous works \citep[e.g., ][]{Krawczyk2013, Ross2015, Bilick2016} found that cross-matches performed with a radius of 2-3'' between the user catalogue and WISE was a good compromise between completeness and contamination \citep[see e.g. Figure 4 of ][]{Krawczyk2013}. We thus use a cross-match radius of 2.75''. When querying the ALLWISE catalogs, ``Cone Search Radius'' in the ALLWISE table search was set to 2.75'' for the Spatial Constraints. The ``One to One Match'' was ``not'' checked; although possible, we consider it highly unlikely there would be more than one MIR source contributing to the flux of a single UV/optically bright rest-frame quasar. Investigating this in detail is very interesting but left to a future study.

@@

First, we will concentrate on detection rate in the infrared, go on to
report on the color-redshift and color-color properties of our sample
and then report on how the current sample populates the
luminosity-redshift $Lz$-plane.

@@

There are 14 objects that have no NIR detections.  3 of these
(PSOJ053.9605-15.7956, PSO J056.7168-16.4769 and DELSJ0411-0907) have
been observed (by VHS) but are out of our queried time range.  6
objects have not been observed (or at least the data is not in the
WSA/VSA archives yet) and 5 objects at declination $\delta \geq +60$
deg are too far north for UKIRT and cannot be observed.

  \subsubsection{Comparing WFCAM and VISTA}
  There are 130 overlapping QSOs between WFCAM and VISTA. Using the
  {\tt VegaToAB} value to put these objects on the same AB system, and
  for each object compared the two measurements. First, the calculated
  weighted average (calibrated flux) in each filter of both and
  calculated the ratio and difference between each measurement and the
  average.  Then for each filter we calculated the weighted average of
  the differences (in mag) for each instrument to see if there were
  significant offsets. The results are given in
  Table~\ref{tab:WFCAM_vs_VISTA}.  The only filter with a significant
  offset is the $Y$-band. All of the VISTA averages are negative and all
  of the WFCAM ones are positive. The $Ks$ versus $K$ band may be
  lightly dodgy, given the different shapes of the filters.
  \begin{table}
    \centering
    \begin{tabular}{l r r}
      \hline  \hline
      abs(VIRCAM & \multirow{2}{*}{millimags} &  no. of  \\
      -  WFCAM)      &                                        &  objects \\
      \hline
      $Z$                 &  23.2 	& 3 \\
      $Y$                 &  57.3 	& 53 \\
      $J$                  &    2.1 	& 106 \\
      $H$                 &  45.8     &  96 \\
      $K_{\rm s}$/$K$ &  25.2     & 110 \\
      \hline  \hline
    \end{tabular}
    \caption{Comparing the magnitudes in different WFCAM/UKIRT and 
      VIRCAM/VISTA near-infrared bands.}
    \label{tab:WFCAM_vs_VISTA}
  \end{table}

@@

\begin{table}
  \centering
  \begin{tabular}{l r r r r}
    \hline \hline
    \multirow{ 2}{*}{$z \geq$} & \multicolumn{2}{c}{unWISE} &  \multicolumn{2}{c}{ALLWISE}   \\ 
                                             & \multicolumn{1}{c}{W1} & \multicolumn{1}{c}{W2} & \multicolumn{1}{c}{W3} & \multicolumn{1}{c}{W4}  \\ 
    \hline
    5.00   &   362 ( 78.2\%)   &   308 ( 66.5\%)  &    51 ( 11.0\%)  &    10 (  2.2\%) \\ 
    5.70   &   186 ( 69.4\%)   &   151 ( 56.3\%)  &    15 (  5.6\%)  &     2 (  0.7\%) \\ 
    6.00   &   109 ( 63.7\%)   &    91 ( 53.2\%)  &     8 (  4.7\%)  &     2 (  1.2\%) \\ 
    6.19   &    63 ( 72.4\%)   &    47 ( 54.0\%)  &     3 (  3.4\%)  &     2 (  2.3\%) \\ 
    6.50   &    34 ( 85.0\%)   &    26 ( 65.0\%)  &     1 (  2.5\%)  &     2 (  5.0\%) \\ 
    6.78   &    12 ( 85.7\%)   &     9 ( 64.3\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    7.00   &     4 (100.0\%)   &     4 (100.0\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    7.50   &     1 (100.0\%)   &     1 (100.0\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    \hline \hline
  \end{tabular}
  \caption{Detection rates in the mid-infrared bands from WISE as a function of 
    redshift. A signal-to-noise cut of {\tt w$x$snr}$>3.0$ has been used for each band.}
  \label{tab:WFCAM_vs_VISTA}
\end{table}

@@

Before reporting on the detection rates, we investigate this effect 
using the AllWISE Source Catalog. 
Figure~\ref{fig:WISEmag_vs_coverage} shows the WISE magnitude versus signal-to-noise, colour coded by {\tt w$x$cov} the mean coverage depth, in each corresponding band. In the two shorter bands W1/2 we see the clear and expected trend for brighter objects to have larger SNR, and also for the higher signal to noise for objects with more exposures at a given magnitude. The behaviour for the W3/4 bands is different, with two populations clearly evident in W3 and although a bit more mixed, also in W4. With the suggested split at SNR$>2$, and no obvious R.A./Declination dependence seen, this behaviour is explained by the fact that there are non-detectiopns in W3/4 for objects (with high W1/2 SNR) that are reported in the ALLWISE catalogue.
AllWISE Source Catalog

For the 278 VH$z$Q with coverage detections, the mean number of exposures for the W1/2 bands is 32.0 and 31.5, respectively, with a minimum number of exposures 17 and 12, and the maximum number or exposures being 114 (for both bands).  For the W3/4 filters, the corresponding mean, minimum and maximum exposure are 17.4 and 17.5, 5.8 and 6.8 and 69 (for both bands). These values are direclty from the {\tt w$x$cov} enteries in the WISE \href {http://wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2_1a.html#w1cov}{ALLWISE catalogue}.

%% detections_vs_WISEcoverage_plots.py
\begin{figure}
  %% trim=l b r t
  \includegraphics[width=8.6cm, clip,trim=6mm 6mm 0mm 6mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/detections/WISEmag_vs_coverage_20190605.pdf}
  \centering
  \vspace{-14pt}
  \caption[]{WISE W1/2/3/4 {\tt w$x$mpro} magnitude against signal-to-noise, 
    colour coded by {\tt w$x$cov} the mean coverage depth, in each corresponding band.
  }
  \label{fig:WISEmag_vs_coverage}
\end{figure}

>>>>

We use data from the the beginning of the WISE mission \citep[2010
January; ][]{Wright2010} through the fifth-year of NEOWISE-R
operations \citep[2018 December;]{Mainzer2011}.  More specifically, we
utilise the recently released ``unWISE Catalog''
\citet{Schlafly2019}. The unWISE
effort\footnote{\href{http://unwise.me}{http://unwise.me}} is the
unblurred coadds of the WISE imaging using the AllWISE and NEOWISE-R
stacked data \citep{Lang2014, Meisner2018a, Meisner2018b}.

Previous works \citep[e.g., ][]{Krawczyk2013, Ross2015, Bilick2016}
found that cross-matches performed with a radius of 2-3'' between the user
catalogue and WISE was a good compromise between completeness and
contamination \citep[see e.g. Figure 4 of ][]{Krawczyk2013}. 
We thus use a cross-match radius of 2.75''.

For the two shorter WISE bands, $\lambda_{\rm eff}$3.37$\mu$m W1 and 
$\lambda_{\rm eff}$4.62$\mu$m W2 we use the deeper, unblurred unWISE coadd data. 
For the two longer WISE bands, $\lambda_{\rm eff}$12.1$\mu$m and $\lambda_{\rm eff}$22.8$\mu$m 
we use the ALLWISE Data Release

When querying the ALLWISE catalogs, ``Cone Search Radius'' in the ALLWISE table search was set to 2.75'' for the Spatial Constraints. The ``One to One Match'' was ``not'' checked; although possible, we consider it highly unlikely there would be more than one MIR source contributing to the flux of a single UV/optically bright rest-frame quasar. Investigating this in detail is very interesting but left to a future study.

@@

First, we will concentrate on detection rate in the infrared, go on to
report on the colour-redshift and colour-colour properties of our sample
and then report on how the current sample populates the
luminosity-redshift $Lz$-plane.

@@

There are 14 objects that have no NIR detections.  3 of these
(PSOJ053.9605-15.7956, PSO J056.7168-16.4769 and DELSJ0411-0907) have
been observed (by VHS) but are out of our queried time range 
(for which the data is publicly available). 6 objects have not been 
observed (or at least the data is not in the
WSA/VSA archives yet) and 5 objects are at a declination $\delta \geq +60$
deg are too far north for UKIRT and cannot be observed.

  \subsubsection{Comparing WFCAM and VISTA}
  There are 130 overlapping QSOs between WFCAM and VISTA. We used the
  {\tt VegaToAB} values to put these objects on the same AB system, and
  for each object compared the two measurements. First, we calculated the
  weighted average (calibrated flux) in each filter of both and
  calculated the ratio and difference between each measurement and the
  average.  Then for each filter we calculated the weighted average of
  the differences (in mag) for each instrument to see if there were
  significant offsets. The results are given in
  Table~\ref{tab:WFCAM_vs_VISTA}.  The only filter with a significant
  offset is the $Y$-band. All of the VISTA averages are negative and all
  of the WFCAM ones are positive. The $Ks$ versus $K$ band comparison may be
  affected the different shapes of the filters, with $K$ being significantly wider than $K_s$.
  \begin{table}
    \centering
    \begin{tabular}{l r r}
      \hline  \hline
      abs(VIRCAM & \multirow{2}{*}{millimags} &  no. of  \\
      -  WFCAM)      &                                        &  objects \\
      \hline
      $Z$                 &  23.2 	& 3 \\
      $Y$                 &  57.3 	& 53 \\
      $J$                  &    2.1 	& 106 \\
      $H$                 &  45.8     &  96 \\
      $K_{\rm s}$/$K$ &  25.2     & 110 \\
      \hline  \hline
    \end{tabular}
    \caption{Comparing the magnitudes in different WFCAM/UKIRT and 
      VIRCAM/VISTA near-infrared bands.}
    \label{tab:WFCAM_vs_VISTA}
  \end{table}

@@

\begin{table}
  \centering
  \begin{tabular}{l r r r r}
    \hline \hline
    \multirow{ 2}{*}{$z \geq$} & \multicolumn{2}{c}{unWISE} &  \multicolumn{2}{c}{ALLWISE}   \\ 
                                             & \multicolumn{1}{c}{W1} & \multicolumn{1}{c}{W2} & \multicolumn{1}{c}{W3} & \multicolumn{1}{c}{W4}  \\ 
    \hline
    5.00   &   362 ( 78.2\%)   &   308 ( 66.5\%)  &    51 ( 11.0\%)  &    10 (  2.2\%) \\ 
    5.70   &   186 ( 69.4\%)   &   151 ( 56.3\%)  &    15 (  5.6\%)  &     2 (  0.7\%) \\ 
    6.00   &   109 ( 63.7\%)   &    91 ( 53.2\%)  &     8 (  4.7\%)  &     2 (  1.2\%) \\ 
    6.19   &    63 ( 72.4\%)   &    47 ( 54.0\%)  &     3 (  3.4\%)  &     2 (  2.3\%) \\ 
    6.50   &    34 ( 85.0\%)   &    26 ( 65.0\%)  &     1 (  2.5\%)  &     2 (  5.0\%) \\ 
    6.78   &    12 ( 85.7\%)   &     9 ( 64.3\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    7.00   &     4 (100.0\%)   &     4 (100.0\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    7.50   &     1 (100.0\%)   &     1 (100.0\%)  &     0 (  0.0\%)  &     0 (  0.0\%) \\ 
    \hline \line
  \end{tabular}
  \caption{Detection rates in the mid-infrared bands from WISE as a function of 
    redshift. A signal-to-noise cut of {\tt w$x$snr}$>3.0$ has been used for each band.}
  \label{tab:WFCAM_vs_VISTA}
\end{table}

@@

Before reporting on the detection rates, we investigate this effect. Figure~\ref{fig:WISEmag_vs_coverage} shows the WISE magnitude versus signal-to-noise, colour coded by {\tt w$x$cov} the mean coverage depth, in each corresponding band. In the two shorter bands W1/2 we see the clear and expected trend for brighter objects to have larger SNR, and also for the higher signal to noise for objects with more exposures at a given magnitude. The behaviour for the W3/4 bands is different, with two populations clearly evident in W3 and although a bit more mixed, also in W4. With the suggested split at SNR$>2$, and no obvious R.A./Declination dependence seen, this behaviour is explained by the fact that there are non-detections in W3/4 for objects (with high W1/2 SNR) that are reported in the ALLWISE catalogue.

For the 278 VH$z$Q with coverage detections, the mean number of exposures for the W1/2 bands is 32.0 and 31.5, respectively, with a minimum number of exposures 17 and 12, and the maximum number or exposures being 114 (for both bands).  For the W3/4 filters, the corresponding mean, minimum and maximum exposure are 17.4 and 17.5, 5.8 and 6.8 and 69 (for both bands). These values are directly from the {\tt w$x$cov} entries in the WISE \href {http://wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2_1a.html#w1cov}{ALLWISE catalogue}.

\begin{figure}
  %% trim=l b r t
  \includegraphics[width=8.6cm, clip,trim=6mm 6mm 0mm 6mm]
  {/cos_pc19a_npr/programs/quasars/highest_z/detections/WISEmag_vs_coverage_2x2_v1.pdf}
  \centering
  \vspace{-14pt}
  \caption[]{WISE W1/2/3/4 magnitude against signal-to-noise, 
    colour coded by w$x$cov the mean coverage depth, in each corresponding band.
  }
  \label{fig:WISEmag_vs_coverage}
\end{figure}
